---
title: "CRC Biomarker Survival Analysis by Treatment MoA"
author: "Medha Rao"
date: today
format: pdf
---

# Preliminary Project 

This project aims to identify genetic biomarkers that are associated with overall survival by various treatment mechanism of action groups amongst colorectal cancer patients. 
Database used: MSK-CHORD

```{r Library Loading, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse) # data handling, plots 
library(gt) # "good table" to make pretty tables
library(survival) # survivial analysis 
library(boot) # boostrapping analyses
library(survminer) # Kaplan-Meier plots 
library(reshape2) # correlation heat map creation helper
```

```{r Read in data, echo = FALSE}
data_timeline_mmr <- read.delim(file = "msk_chord_2024/data_timeline_mmr.txt", header = TRUE)

data_timeline_pdl1<- read.delim(file = "msk_chord_2024/data_timeline_pdl1.txt", header = TRUE)

data_timeline_prior_meds <- read.delim(file = "msk_chord_2024/data_timeline_prior_meds.txt", header = TRUE)

data_timeline_radiation <- read.delim(file = "msk_chord_2024/data_timeline_radiation.txt", header = TRUE)

data_timeline_surgery <- read.delim(file = "msk_chord_2024/data_timeline_surgery.txt", header = TRUE)

data_clinical_patient <- read.delim(file = "msk_chord_2024/data_clinical_patient.txt", header = TRUE, skip = 4)

data_clinical_sample <- read.delim(file = "msk_chord_2024/data_clinical_sample.txt", header = TRUE, skip = 4)

data_timeline_diagnosis <- read.delim(file = "msk_chord_2024/data_timeline_diagnosis.txt", header = TRUE)

data_timeline_treatment <- read.delim(file = "msk_chord_2024/data_timeline_treatment.txt", header = TRUE)

data_mutations <- read.delim(file = "msk_chord_2024/data_mutations.txt", header = TRUE)
```

```{r CRC only data sets}
# Identifying only CRC patients
CRC_patients <- data_clinical_sample |> filter(CANCER_TYPE == "Colorectal Cancer") |>
  pull(PATIENT_ID) # CLINICAL SAMPLE has MSI flags
  # 5543 CRC patients

# Consolidation of clinical data sets (demographic info, cancer info, MSI info, etc.)
CRC_clinical_pt <- data_clinical_patient |> filter(PATIENT_ID %in% CRC_patients) 
CRC_clinical_samp<- data_clinical_sample |> filter(PATIENT_ID %in% CRC_patients)

CRC_clinical <- full_join(CRC_clinical_pt, CRC_clinical_samp)

# timeline data --> only selecting CRC patients
CRC_diagnosis <- data_timeline_diagnosis |> filter(PATIENT_ID %in% CRC_patients)

CRC_mmr <- data_timeline_mmr |> filter(PATIENT_ID %in% CRC_patients)

CRC_pdl1 <- data_timeline_pdl1 |> filter(PATIENT_ID %in% CRC_patients)

CRC_priorMeds <- data_timeline_prior_meds |> filter(PATIENT_ID %in% CRC_patients)

CRC_radiation <- data_timeline_radiation |> filter(PATIENT_ID %in% CRC_patients)

CRC_surgery <- data_timeline_surgery |> filter(PATIENT_ID %in% CRC_patients)

CRC_tx <- data_timeline_treatment |> filter(PATIENT_ID %in% CRC_patients)

CRC_muts <- data_mutations |> 
  mutate(PATIENT_ID = str_extract(Tumor_Sample_Barcode, "P-\\d{7}")) |>
  filter(PATIENT_ID %in% CRC_patients)

# cleaning up environment :) 
rm(data_clinical_patient, data_clinical_sample, data_mutations, data_timeline_diagnosis, data_timeline_mmr, data_timeline_pdl1, data_timeline_prior_meds, data_timeline_radiation, data_timeline_surgery, data_timeline_treatment, CRC_clinical_pt, CRC_clinical_samp)
```

```{r Drug/ MoA mapping }
# note: MoA is the same as saying treatment mechanism of action 

# create a mapping from each possible ( from Feb 2026 available info) CRC drug name to a drug treatment MoA category
drug_moa <- tibble::tribble(
  ~drug, ~moa,
  "VINCRISTINE", "CHEMO_CYTOTOXIC",
  "FLUDARABINE", "CHEMO_CYTOTOXIC",
  "CYCLOPHOSPHAMIDE", "CHEMO_CYTOTOXIC",
  "CISPLATIN", "CHEMO_CYTOTOXIC",
  "GEMCITABINE", "CHEMO_CYTOTOXIC",
  "CAPECITABINE", "CHEMO_CYTOTOXIC",
  "FLUOROURACIL", "CHEMO_CYTOTOXIC",
  "LEUCOVORIN", "CHEMO_CYTOTOXIC",
  "OXALIPLATIN", "CHEMO_CYTOTOXIC",
  "IRINOTECAN", "CHEMO_CYTOTOXIC",
  "BEVACIZUMAB", "ANTI_VEGF_ANGIO",
  "PANITUMUMAB", "ANTI_EGFR",
  "INVESTIGATIONAL", "UNKNOWN_INVESTIGATIONAL",
  "ZOLEDRONIC ACID", "HORMONAL_SUPPORTIVE_OTHER",
  "FLOXURIDINE", "CHEMO_CYTOTOXIC",
  "VEMURAFENIB", "TARGETED_OTHER",
  "CETUXIMAB", "ANTI_EGFR",
  "MEGESTROL", "HORMONAL_SUPPORTIVE_OTHER",
  "TIPIRACIL-TRIFLURIDINE", "CHEMO_CYTOTOXIC",
  "REGORAFENIB", "ANTI_VEGF_ANGIO",
  "TRASTUZUMAB", "TARGETED_OTHER",
  "YTTRIUM Y-90 THERASPHERES", "RADIOPHARM",
  "MITOMYCIN", "CHEMO_CYTOTOXIC",
  "PAMIDRONATE", "HORMONAL_SUPPORTIVE_OTHER",
  "DABRAFENIB", "TARGETED_OTHER",
  "TRAMETINIB", "TARGETED_OTHER",
  "CARBOPLATIN", "CHEMO_CYTOTOXIC",
  "ETOPOSIDE", "CHEMO_CYTOTOXIC",
  "COBIMETINIB", "TARGETED_OTHER",
  "ISOTRETINOIN", "HORMONAL_SUPPORTIVE_OTHER",
  "RITUXIMAB", "IMMUNOTHERAPY",
  "SODIUM IODIDE I-131", "RADIOPHARM",
  "PEMBROLIZUMAB", "IMMUNOTHERAPY",
  "YTTRIUM Y-90 MICROSPHERES", "RADIOPHARM",
  "IPILIMUMAB", "IMMUNOTHERAPY",
  "NIVOLUMAB", "IMMUNOTHERAPY",
  "METHOTREXATE", "CHEMO_CYTOTOXIC",
  "EVEROLIMUS", "TARGETED_OTHER",
  "FLUOROURACIL TOPICAL", "CHEMO_CYTOTOXIC",
  "ANASTROZOLE", "HORMONAL_SUPPORTIVE_OTHER",
  "CLADRIBINE", "CHEMO_CYTOTOXIC",
  "LETROZOLE", "HORMONAL_SUPPORTIVE_OTHER",
  "BICALUTAMIDE", "HORMONAL_SUPPORTIVE_OTHER",
  "DEGARELIX", "HORMONAL_SUPPORTIVE_OTHER",
  "RALOXIFENE", "HORMONAL_SUPPORTIVE_OTHER",
  "IMATINIB", "TARGETED_OTHER",
  "SUNITINIB", "ANTI_VEGF_ANGIO",
  "SODIUM IODIDE I-131 MIBG", "RADIOPHARM",
  "LUTETIUM LU-177 DOTATATE", "RADIOPHARM",
  "DOXORUBICIN", "CHEMO_CYTOTOXIC",
  "PACLITAXEL", "CHEMO_CYTOTOXIC",
  "CYTARABINE", "CHEMO_CYTOTOXIC",
  "CARMUSTINE", "CHEMO_CYTOTOXIC",
  "MELPHALAN", "CHEMO_CYTOTOXIC",
  "EXEMESTANE", "HORMONAL_SUPPORTIVE_OTHER",
  "ATEZOLIZUMAB", "IMMUNOTHERAPY",
  "HYDROXYUREA", "CHEMO_CYTOTOXIC",
  "AZACITIDINE", "CHEMO_CYTOTOXIC",
  "LEUPROLIDE", "HORMONAL_SUPPORTIVE_OTHER",
  "ENCORAFENIB", "TARGETED_OTHER",
  "BLEOMYCIN", "CHEMO_CYTOTOXIC",
  "DACTINOMYCIN", "CHEMO_CYTOTOXIC",
  "IFOSFAMIDE", "CHEMO_CYTOTOXIC",
  "TEMOZOLOMIDE", "CHEMO_CYTOTOXIC",
  "FULVESTRANT", "HORMONAL_SUPPORTIVE_OTHER",
  "PALBOCICLIB", "TARGETED_OTHER",
  "TAMOXIFEN", "HORMONAL_SUPPORTIVE_OTHER",
  "PAZOPANIB", "ANTI_VEGF_ANGIO",
  "PACLITAXEL PROTEIN-BOUND", "CHEMO_CYTOTOXIC",
  "RAMUCIRUMAB", "ANTI_VEGF_ANGIO",
  "MERCAPTOPURINE", "CHEMO_CYTOTOXIC",
  "RUXOLITINIB", "TARGETED_OTHER",
  "PACRITINIB", "TARGETED_OTHER",
  "GOSERELIN", "HORMONAL_SUPPORTIVE_OTHER",
  "PROCARBAZINE", "CHEMO_CYTOTOXIC",
  "LENALIDOMIDE", "TARGETED_OTHER",
  "DOCETAXEL", "CHEMO_CYTOTOXIC",
  "ABIRATERONE", "HORMONAL_SUPPORTIVE_OTHER",
  "APALUTAMIDE", "HORMONAL_SUPPORTIVE_OTHER",
  "CABAZITAXEL", "CHEMO_CYTOTOXIC",
  "OLAPARIB", "TARGETED_OTHER",
  "TUCATINIB", "TARGETED_OTHER",
  "ASPARAGINASE ESCHERICHIA COLI", "CHEMO_CYTOTOXIC",
  "PEGASPARGASE", "CHEMO_CYTOTOXIC",
  "ALEMTUZUMAB", "IMMUNOTHERAPY",
  "PEMETREXED", "CHEMO_CYTOTOXIC",
  "ADO-TRASTUZUMAB EMTANSINE", "TARGETED_OTHER",
  "PERTUZUMAB", "TARGETED_OTHER",
  "BRENTUXIMAB VEDOTIN", "IMMUNOTHERAPY",
  "MITOXANTRONE", "CHEMO_CYTOTOXIC",
  "OBINUTUZUMAB", "IMMUNOTHERAPY",
  "ENZALUTAMIDE", "HORMONAL_SUPPORTIVE_OTHER",
  "TALAZOPARIB", "TARGETED_OTHER",
  "ERLOTINIB", "TARGETED_OTHER",
  "IBRUTINIB", "TARGETED_OTHER",
  "DAROLUTAMIDE", "HORMONAL_SUPPORTIVE_OTHER",
  "LAPATINIB", "TARGETED_OTHER",
  "BCG", "HORMONAL_SUPPORTIVE_OTHER",
  "LENVATINIB", "ANTI_VEGF_ANGIO",
  "BINIMETINIB", "TARGETED_OTHER",
  "FAM-TRASTUZUMAB DERUXTECAN", "TARGETED_OTHER",
  "DOXORUBICIN LIPOSOMAL", "CHEMO_CYTOTOXIC",
  "VINORELBINE", "CHEMO_CYTOTOXIC",
  "RADIUM RA-223 DICHLORIDE", "RADIOPHARM",
  "TRETINOIN", "HORMONAL_SUPPORTIVE_OTHER",
  "NIRAPARIB", "TARGETED_OTHER",
  "CARFILZOMIB", "TARGETED_OTHER",
  "CABOZANTINIB", "ANTI_VEGF_ANGIO",
  "SOTORASIB", "TARGETED_OTHER",
  "CERITINIB", "TARGETED_OTHER",
  "RIBOCICLIB", "TARGETED_OTHER",
  "ACALABRUTINIB", "TARGETED_OTHER",
  "THIOTEPA", "CHEMO_CYTOTOXIC",
  "BENDAMUSTINE", "CHEMO_CYTOTOXIC",
  "CRIZOTINIB", "TARGETED_OTHER",
  "VISMODEGIB", "TARGETED_OTHER",
  "AFLIBERCEPT OPHTHALMIC", "ANTI_VEGF_ANGIO",
  "SELPERCATINIB", "TARGETED_OTHER",
  "NERATINIB", "TARGETED_OTHER",
  "VENETOCLAX", "TARGETED_OTHER",
  "DARATUMUMAB", "IMMUNOTHERAPY",
  "POMALIDOMIDE", "TARGETED_OTHER",
  "IRINOTECAN LIPOSOMAL", "CHEMO_CYTOTOXIC",
  "DASATINIB", "TARGETED_OTHER",
  "PONATINIB", "TARGETED_OTHER",
  "BLINATUMOMAB", "IMMUNOTHERAPY",
  "LAROTRECTINIB", "TARGETED_OTHER",
  "DURVALUMAB", "IMMUNOTHERAPY",
  "NILOTINIB", "TARGETED_OTHER",
  "AVELUMAB", "IMMUNOTHERAPY",
  "ENFORTUMAB VEDOTIN", "IMMUNOTHERAPY",
  "RUCAPARIB", "TARGETED_OTHER",
  "DAUNORUBICIN", "CHEMO_CYTOTOXIC",
  "GEMTUZUMAB OZOGAMICIN", "IMMUNOTHERAPY",
  "BUSULFAN", "CHEMO_CYTOTOXIC",
  "CLOFARABINE", "CHEMO_CYTOTOXIC",
  "THIOGUANINE", "CHEMO_CYTOTOXIC",
  "IVOSIDENIB", "TARGETED_OTHER",
  "PENTOSTATIN", "CHEMO_CYTOTOXIC",
  "SORAFENIB", "ANTI_VEGF_ANGIO",
  "BORTEZOMIB", "TARGETED_OTHER",
  "DARATUMUMAB-HYALURONIDASE", "IMMUNOTHERAPY",
  "HYALURONIDASE-RITUXIMAB", "IMMUNOTHERAPY",
  "BRIGATINIB", "TARGETED_OTHER",
  "IXAZOMIB", "TARGETED_OTHER",
  "EPIRUBICIN", "CHEMO_CYTOTOXIC",
  "ROMIDEPSIN", "CHEMO_CYTOTOXIC",
  "TOPOTECAN", "CHEMO_CYTOTOXIC",
  "VINBLASTINE", "CHEMO_CYTOTOXIC",
  "OFATUMUMAB", "IMMUNOTHERAPY",
  "ENTRECTINIB", "TARGETED_OTHER",
  "POLATUZUMAB VEDOTIN", "IMMUNOTHERAPY",
  "EPCORITAMAB", "IMMUNOTHERAPY"
)

# join the treatments to the initial drug mapping
CRC_tx <- CRC_tx |> 
  left_join(drug_moa, by = c("AGENT" = "drug"))
```

```{r Create survival cohort}
# only doing metastatic CRC, makes it easier to control for differential treatment assignments 

metaCRC_patients <- CRC_clinical |> 
  filter(METASTATIC_SITE != "Not Applicable") |> 
  dplyr::select(PATIENT_ID) |> 
  dplyr::pull()

Survival_patients <- CRC_tx |> 
  filter(PATIENT_ID %in% metaCRC_patients) |> 
  dplyr::select(PATIENT_ID) |> 
  dplyr::pull()

CRC_tx <- CRC_tx |> 
  filter(PATIENT_ID %in% metaCRC_patients)

CRC_diagnosis <- CRC_diagnosis |> 
  filter(PATIENT_ID %in% Survival_patients)

# find duplicated rows, some have same CRC diagnosis split into different parts of intestine, some also have lung cancer in this data set, need to take the lung cancer diagnoses out since that date is not important

dupe_rows <- CRC_diagnosis |>
  group_by(PATIENT_ID) |>
  filter(n() > 1) |>
  ungroup() |> 
  dplyr::select(PATIENT_ID, START_DATE, DX_DESCRIPTION) |>
  mutate(CRC_yes = case_when(str_detect(DX_DESCRIPTION, "COLON|RECTUM|RECTOSIGMOID") ~ 1,
                             TRUE ~ 0)) |> 
  filter(CRC_yes == 1) |> 
  group_by(PATIENT_ID) |>
  mutate(keepThis = case_when(
    n() == 1 ~ TRUE,
    n() > 1 & min(START_DATE) == max(START_DATE) & row_number() == 1 ~ TRUE,
    n() > 1 & min(START_DATE) != max(START_DATE) ~ TRUE, 
    TRUE ~ FALSE
  )) |> 
  ungroup() |> dplyr::select(-CRC_yes)

# final CRC diagnosis set, now only has relevant CRC diagnosis (metastatic)
CRC_diagnosis <- CRC_diagnosis |> 
  left_join(dupe_rows, by = c("PATIENT_ID", "START_DATE", "DX_DESCRIPTION")) |> 
  filter(keepThis == TRUE | is.na(keepThis)) |> dplyr::select(-keepThis)

# now need to identify the unique start dates of the patient diagnosis (same diagnosis of multiple areas on the same day count as one diagnosis of CRC on that day)
dx_all_days_per_patient <- CRC_diagnosis |>
  group_by(PATIENT_ID) |>
  summarise(
    dx_days_from_seq = list(sort(unique(START_DATE))),
    n_dx = n_distinct(START_DATE),
    .groups = "drop"
  )

# using the diagnosis and start dates of diagnosis and joining into the treatment data set 
try_tx_filtering <- left_join(CRC_tx, dx_all_days_per_patient, by = "PATIENT_ID")

# patients have many treatments at different time intervals, this chooses the proper diagnosis date that corresponds with the proper firstline treatment. For example, they could have been treated for lung cancer two years ago, but that counts as the first line unless it is filtered out 
dx_chosen <- try_tx_filtering |>
  distinct(PATIENT_ID, dx_days_from_seq, START_DATE) |>
  group_by(PATIENT_ID) |>
  summarise(
    first_tx_day = suppressWarnings(min(as.numeric(START_DATE), na.rm = TRUE)),
    dx_list = list(
      as.numeric(str_trim(str_extract_all(dx_days_from_seq, "-?\\d+")[[1]]))
    ),
    dx_day = {
      d <- dx_list[[1]]
      d_ok <- d[!is.na(d) & d <= first_tx_day]
      if (length(d_ok) == 0) NA_real_ else max(d_ok)
    },
    .groups = "drop"
  )

# choose the firstline treatment as the earliest treatment ON or AFTER the diagnosis date  
firstline_tx <- try_tx_filtering |>
  left_join(dx_chosen |> dplyr::select(PATIENT_ID, dx_day), by = "PATIENT_ID") |>
  mutate(START_DATE = as.numeric(START_DATE)) |>
  filter(!is.na(dx_day)) |>
  group_by(PATIENT_ID) |>
  filter(START_DATE >= dx_day) |>
  mutate(first1l_day = min(START_DATE, na.rm = TRUE)) |>
  filter(START_DATE == first1l_day) |>
  ungroup()

# summary of the regimen started on firstline treatment (i.e. started chemo and a targeted agent in tandem)
firstline_regimen <- firstline_tx |>
  group_by(PATIENT_ID) |>
  summarise(
    dx_day = first(dx_day),
    first1l_day = first(first1l_day),
    agents = paste(sort(unique(AGENT)), collapse = " + "),
    moas   = paste(sort(unique(moa)), collapse = " + "),
    .groups = "drop"
  )

# joining clinical and firstline regimen data 
CRC_clinical <- left_join(firstline_regimen, CRC_clinical, by = "PATIENT_ID")

# final making sure that it is the metastatic patients and only CRC-related rows 
CRC_clinical <- CRC_clinical |> filter(CANCER_TYPE == "Colorectal Cancer") |> filter (PATIENT_ID %in% metaCRC_patients)

# Create moa_grp from moas combinations
CRC_clinical <- CRC_clinical |>
  mutate(moa_grp = case_when(
    moas == "CHEMO_CYTOTOXIC" ~ "CHEMO",
    moas == "ANTI_VEGF_ANGIO + CHEMO_CYTOTOXIC" ~ "CHEMO+VEGF",
    moas == "ANTI_EGFR + CHEMO_CYTOTOXIC" ~ "CHEMO+EGFR",
    moas %in% c("IMMUNOTHERAPY") ~ "IO",
    TRUE ~ "OTHER"
  ))

CRC_clinical <- CRC_clinical |>
  mutate(moa_consolidated = case_when(
    moa_grp == "CHEMO" ~ "Chemo",
    moa_grp %in% c("CHEMO+VEGF", "CHEMO+EGFR") ~ "Chemo+Targeted",
    moa_grp %in% c("IO", "OTHER") ~ "Other/IO",
    TRUE ~ NA_character_
  ))

table(CRC_clinical$moa_consolidated)
```

```{r Top genes and mutations}
# Get mutation frequency for all genes
mutationFreq <- CRC_muts |>
     group_by(Hugo_Symbol) |>
     summarise(nPats = n_distinct(PATIENT_ID), .groups = "drop") |>
     arrange(desc(nPats))

head(mutationFreq, 15)

# Choose top 10 genes (chose 10 arbitrarily, can choose a different number too)
topGenes <- mutationFreq |>
     head(10) |>
     pull(Hugo_Symbol)

print(topGenes)

mutationGenes <- topGenes

# matrix of mutations by patient
mutMat <- CRC_muts |>
  distinct(PATIENT_ID, Hugo_Symbol) |>
  pivot_wider(
    names_from = Hugo_Symbol,
    values_from = Hugo_Symbol,
    values_fill = NA
  ) |>
  # Convert to binary (1 = mutation)
  mutate(across(all_of(mutationGenes),
                ~ifelse(!is.na(.), 1, 0), 
                .names = "{.col}_mut"))

# Join mutations to clinical data
survDF<- CRC_clinical |>
  left_join(mutMat |> dplyr::select(PATIENT_ID, ends_with("_mut")), 
            by = "PATIENT_ID") |>
  # Set NA mutations to 0 
  mutate(across(ends_with("_mut"), 
                ~replace_na(., 0))) |>
  # Create survival event indicator
  mutate(
    OS_event = as.numeric(OS_STATUS == "1:DECEASED"),
    OS_time = OS_MONTHS - (first1l_day / 30.44) # time 0 is start of first 1L regimen! 
  ) |>
  # Remove rows with missing time or event data
  filter(!is.na(OS_time) & !is.na(OS_event) & OS_time > 0)

# check for mutation prevalence in cohort
survDF |>
  summarise(across(ends_with("_mut"), 
                   list(n = ~sum(., na.rm = TRUE),
                        pct = ~mean(., na.rm = TRUE) * 100),
                   .names = "{.col}_{.fn}")) |>
  t() |>
  print()

# Add transformed variables, done after linearity assumption check 
survDF <- survDF |> 
  mutate(ageSquared = CURRENT_AGE_DEID^2, 
                           logMSI = log(MSI_SCORE + 1), 
                           logTMB = log(TMB_NONSYNONYMOUS + 1), 
                           logMSISquared = logMSI^2, 
                           logTMBSquared = logTMB^2) |> 
  filter(!is.na(MSI_SCORE)) |>
  filter(MSI_SCORE != -1)
```

```{r Simple Cox Univariate Model}

# create a list of results for the univar models I am running
uniRes <- list()

# for loop to build a model for each gene in top N gene list (choosing top 10 for now)
for (gene in mutationGenes) {
  geneVar <- paste0(gene, "_mut")
   
  modelFormula <- as.formula(paste(
    "Surv(OS_time, OS_event) ~", geneVar, 
    "+ moa_consolidated +", geneVar, ":moa_consolidated",
    "+ CURRENT_AGE_DEID + ageSquared + GENDER + RACE + STAGE_HIGHEST_RECORDED"
  ))   
  
  cox <- coxph(modelFormula, data = survDF)
  
  uniRes[[gene]] <- list(
    model = cox,
    summary = summary(cox)
  )
}
```

```{r Improved Univariate Cox Model}
# improved (hopefully) univar res list with MSI and TMB 
uniRes_improved <- list()

for (gene in mutationGenes) {
  geneVar <- paste0(gene, "_mut")
  
  modelFormula <- as.formula(paste(
    "Surv(OS_time, OS_event) ~", geneVar, 
    "+ moa_consolidated +", geneVar, ":moa_consolidated",
    "+ CURRENT_AGE_DEID + ageSquared + GENDER + RACE + STAGE_HIGHEST_RECORDED",
    " + logMSI + logTMB + logMSISquared + logTMBSquared" 
  ))
  
  cox <- coxph(modelFormula, data = survDF)
  
  uniRes_improved[[gene]] <- list(
    model = cox,
    summary = summary(cox)
  )
}
```

```{r Stratified Analysis }
# create groups array and results list
moa_groups <- unique(survDF$moa_consolidated)
stratRes <- list()

# for loop creating each model by moa group
for (moa in moa_groups) {
  survDF_subset <- survDF |> filter(moa_consolidated == moa)
  
  stratGeneRes <- list()
  
  for (gene in mutationGenes) {
    geneVar <- paste0(gene, "_mut")
    
    # Model: gene + age + sex + stage + MSI + TMB
    modelForm <- as.formula(paste(
      "Surv(OS_time, OS_event) ~", geneVar,
      "+ CURRENT_AGE_DEID + ageSquared + GENDER + RACE + STAGE_HIGHEST_RECORDED  + logMSI + logTMB + logMSISquared + logTMBSquared"
    ))
    
    cox <- coxph(modelForm, data = survDF_subset)
    
    stratGeneRes[[gene]] <- summary(cox)
    
    # get HR for each moa by mutation
    cat("Gene:", gene, "| HR:", 
        round(summary(cox)$conf.int[1, 1], 3), 
        "| p-value:", 
        round(summary(cox)$coefficients[1, 5], 4), "\n")
  }
  
  stratRes[[moa]] <- stratGeneRes
}
```

```{r Interaction Terms Analysis}

ixRes <- tibble()

for (gene in mutationGenes) {
  coxSumm <- uniRes_improved[[gene]]$summary
  
  # get covar results
  coef_df <- as.data.frame(coxSumm$coefficients) |>
    rownames_to_column("Variable") |>
    as_tibble() |>
    mutate(Gene = gene) |>
    # only choose the interaction terms
    filter(str_detect(Variable, ":moa_consolidated"))
  
  ixRes <- bind_rows(ixRes, coef_df)
}

# create ix table
ixTab <- ixRes |>
  dplyr::select(Gene, Variable, coef, `exp(coef)`, `se(coef)`, `z`, `Pr(>|z|)`) |>
  rename(
    HR = `exp(coef)`,
    SE = `se(coef)`,
    Z_stat = `z`,
    P_value = `Pr(>|z|)`,
    Interaction_Term = Variable
  ) |>
  mutate(
    CI_lower = exp(coef - 1.645 * SE),
    CI_upper = exp(coef + 1.645 * SE),
    Significant = ifelse(P_value < 0.1, "YES", "NO"),
    # added interpretaton for ix term
    Interpretation = case_when(
      P_value < 0.1 & HR > 1 ~ "Mutation effect AMPLIFIED in this treatment",
      P_value < 0.1 & HR < 1 ~ "Mutation effect ATTENUATED in this treatment",
      P_value >= 0.1 ~ "No significant interaction"
    )
  ) |>
  dplyr::select(Gene, Interaction_Term, HR, CI_lower, CI_upper, P_value, Significant, Interpretation) |>
  arrange(P_value)

# FDR correction
ixTab <- ixTab |>
  mutate(
    # FDR correction
    P_value_fdr = p.adjust(P_value, method = "BH"),
    
    # Flag significance for uncorrected and corrected (FDR)
    Sig_Uncorrected = ifelse(P_value < 0.05, "YES", "NO"),
    Sig_FDR = ifelse(P_value_fdr < 0.05, "YES", "NO")
  ) |>
  dplyr::select(
    Gene, Interaction_Term, HR, CI_lower, CI_upper,
    P_value, P_value_fdr,
    Sig_Uncorrected, Sig_FDR,
    Interpretation
  ) |>
  arrange(P_value)

print(ixTab)

cat("Total tests:", nrow(ixTab), "\n")
cat("Significant at uncorrected p < 0.05:", sum(ixTab$Sig_Uncorrected == "YES"), "\n")
cat("Significant at FDR p < 0.05:", sum(ixTab$Sig_FDR == "YES"), "\n\n")

# Show which remain robust after correction
robust_findings <- ixTab |>
  filter(Sig_FDR == "YES") |>
  dplyr::select(Gene, Interaction_Term, HR, P_value, P_value_fdr)

if (nrow(robust_findings) > 0) {
  cat("ROBUST FINDINGS (significant after FDR correction):\n\n")
  print(robust_findings)
} else {
  cat("No findings remain significant after FDR correction.\n")
  nominal <- ixTab |>
    filter(Sig_Uncorrected == "YES") |>
    dplyr::select(Gene, Interaction_Term, HR, P_value, P_value_fdr)
  if (nrow(nominal) > 0) {
    print(nominal)
  } else {
    cat("No nominally significant findings (p < 0.05).\n")
  }
}

cat("\n")

sigIX <- ixTab |>
  filter(Sig_Uncorrected == "YES") |>
  arrange(P_value)

if (nrow(sigIX) > 0) {
  cat("Uncorrected significant interactions (p < 0.05):\n")
  print(sigIX |> dplyr::select(Gene, HR, CI_lower, CI_upper, P_value, P_value_fdr))
} else {
  cat("No uncorrected significant interactions (p < 0.05)\n")
}
```

```{r Stratified HR Results}

stratHR <- tibble()

# for loop to go through each gene and get the HRs and p-vals 
for (treatment in names(stratRes)) {
  for (gene in names(stratRes[[treatment]])) {
    
    summary_obj <- stratRes[[treatment]][[gene]]
    
    # Extract HR and p-value
    hr <- summary_obj$conf.int[1, 1]
    hr_lower <- summary_obj$conf.int[1, 3]
    hr_upper <- summary_obj$conf.int[1, 4]
    p_val <- summary_obj$coefficients[1, 5]
    
    stratHR <- bind_rows(
      stratHR,
      tibble(
        Gene = gene,
        Treatment = treatment,
        HR = hr,
        CI_lower = hr_lower,
        CI_upper = hr_upper,
        P_value = p_val
      )
    )
  }
}

# pivot wider, treatments now next to each other 
stratified_wide <- stratHR |>
  pivot_wider(
    names_from = Treatment,
    values_from = c(HR, CI_lower, CI_upper, P_value),
    names_glue = "{Treatment}_{.value}"
  ) |>
  dplyr::select(Gene, starts_with("Chemo_"), starts_with("Chemo+"), starts_with("Other"))

print(stratified_wide)

# genes with largest HR diff by moa group
hr_comparison <- stratHR |>
  dplyr::select(Gene, Treatment, HR) |>
  pivot_wider(
    names_from = Treatment,
    values_from = HR
  ) |>
  mutate(
    Chemo_vs_ChemoTargeted = ifelse(!is.na(Chemo) & !is.na(`Chemo+Targeted`),
                                     abs(log(Chemo) - log(`Chemo+Targeted`)), 
                                     NA),
    Chemo_vs_OtherIO = ifelse(!is.na(Chemo) & !is.na(`Other/IO`),
                              abs(log(Chemo) - log(`Other/IO`)), 
                              NA),
    ChemoTargeted_vs_OtherIO = ifelse(!is.na(`Chemo+Targeted`) & !is.na(`Other/IO`),
                                      abs(log(`Chemo+Targeted`) - log(`Other/IO`)), 
                                      NA)
  ) |>
  arrange(desc(Chemo_vs_ChemoTargeted), desc(Chemo_vs_OtherIO))

cat("Genes with largest HR differences across treatment groups:\n\n")
print(hr_comparison)
```

```{r Prop. Hazards Assumptions Testing}

# Test Prop. Haz assumption on a few models, chose top significant genes 
test_genes <- mutationGenes

ph_results <- tibble()

for (gene in test_genes) {
  # get univar model with interaction
  model <- uniRes_improved[[gene]]$model
  
  # test Prop. Haz assumption
  ph_test <- cox.zph(model)
  
  cat("Gene:", gene, "\n")
  print(ph_test)
  cat("\n")
  
  # Extract p-values and store
  ph_results <- bind_rows(
    ph_results,
    tibble(
      Gene = gene,
      Variable = rownames(ph_test$table),
      ChiSq = ph_test$table[, "chisq"],
      P_value = ph_test$table[, "p"]
    )
  )
}
```

```{r Stratified by Time (early v later follow-up) Analysis}

# Landmark analysis at median follow-up

median_fu <- median(survDF$OS_time, na.rm = TRUE)

# Early period: time 0 to median
survDF_early <- survDF |>
  filter(OS_time <= median_fu) |>
  mutate(OS_time_early = OS_time)

# Late period: patients alive at median
survDF_late <- survDF |>
  filter(OS_time > median_fu) |>
  mutate(
    OS_time_late = OS_time - median_fu,
    OS_event_late = OS_event
  )

# Initialize results tibble
landmark_results <- tibble()

# Fit separate models for early and late periods
for (gene in c(mutationGenes )) {
  geneVar <- paste0(gene, "_mut")
  
  # EARLY
  cox_early <- coxph(
    as.formula(paste(
      "Surv(OS_time_early, OS_event) ~", geneVar, 
      "+ moa_consolidated +", geneVar, ":moa_consolidated",
      "+ CURRENT_AGE_DEID + ageSquared + GENDER + RACE + STAGE_HIGHEST_RECORDED",
      "+ logMSI + logTMB + logMSISquared + logTMBSquared"
    )),
    data = survDF_early
  )
  
  # LATE
  cox_late <- coxph(
    as.formula(paste(
      "Surv(OS_time_late, OS_event_late) ~", geneVar, 
      "+ moa_consolidated +", geneVar, ":moa_consolidated",
      "+ CURRENT_AGE_DEID + ageSquared + GENDER + RACE + STAGE_HIGHEST_RECORDED",
      "+ logMSI + logTMB + logMSISquared + logTMBSquared"
    )),
    data = survDF_late
  )
  
  # Extract interaction coefficients for early period
  early_summary <- as.data.frame(summary(cox_early)$coefficients) |>
    rownames_to_column("Variable") |>
    filter(str_detect(Variable, ":moa_consolidated")) |>
    mutate(
      Gene = gene,
      Period = "Early",
      HR = `exp(coef)`,
      SE = `se(coef)`,
      P_value = `Pr(>|z|)`,
      CI_lower = exp(coef - 1.96 * SE),
      CI_upper = exp(coef + 1.96 * SE)
    ) |>
    select(Gene, Period, Variable, HR, CI_lower, CI_upper, P_value)
  
  # Extract interaction coefficients for late period
  late_summary <- as.data.frame(summary(cox_late)$coefficients) |>
    rownames_to_column("Variable") |>
    filter(str_detect(Variable, ":moa_consolidated")) |>
    mutate(
      Gene = gene,
      Period = "Late",
      HR = `exp(coef)`,
      SE = `se(coef)`,
      P_value = `Pr(>|z|)`,
      CI_lower = exp(coef - 1.96 * SE),
      CI_upper = exp(coef + 1.96 * SE)
    ) |>
    select(Gene, Period, Variable, HR, CI_lower, CI_upper, P_value)
  
  # Bind to results
  landmark_results <- bind_rows(
    landmark_results,
    early_summary,
    late_summary
  )
}

# Apply FDR correction to these 4 tests
landmark_results <- landmark_results |>
  mutate(
    P_value_fdr = p.adjust(P_value, method = "BH"),
    Significant = ifelse(P_value_fdr < 0.05, "YES", "NO")
  ) |>
  arrange(P_value_fdr)


cat("Median follow-up:", round(median_fu, 1), "months\n")
cat("FDR correction applied overall\n\n")

print(landmark_results |>
  select(Gene, Period, Variable, HR, CI_lower, CI_upper, P_value, P_value_fdr, Significant))

# Display as gt table for manuscript
landmark_results |>
  mutate(
    Treatment = str_remove(str_remove(Variable, ".*:moa_consolidated"), "_.*"),
    `95% CI` = paste0("(", round(CI_lower, 2), "-", round(CI_upper, 2), ")")
  ) |>
  select(Gene, Period, Treatment, HR, `95% CI`, P_value, P_value_fdr, Significant) |>
  gt() |>
  tab_header(
    title = "Landmark Analysis: Gene-Treatment Interactions by Follow-Up Period",
    subtitle = paste0("Early: 0-", round(median_fu, 1), " months | Late: >", round(median_fu, 1), " months")
  ) |>
  fmt_number(columns = c(HR), decimals = 2) |>
  fmt_number(columns = c(P_value, P_value_fdr), decimals = 4)


```

```{r C-Index}
# C-index for all univar results from the improved univar model
c_indices <- tibble()

for (gene in mutationGenes) {
  model <- uniRes_improved[[gene]]$model
  summary_mod <- summary(model)
  
  c_index <- summary_mod$concordance[1]
  c_se <- summary_mod$concordance[2]
  
  c_indices <- bind_rows(
    c_indices,
    tibble(
      Gene = gene,
      C_Index = round(c_index, 3),
      SE = round(c_se, 3)
    )
  )
}

print(c_indices)

cat("\nMean C-index across genes:", round(mean(c_indices$C_Index), 3), "\n")
```

```{r Linearity Assumption}
# choosing the BRAF gene's model to check the assumption 
model_adjusted <- uniRes_improved[["BRAF"]]$model

# residuals
mart_resid <- residuals(model_adjusted, type = "martingale")
model_frame <- model.frame(model_adjusted)

par(mfrow = c(1, 3))

age_numeric <- model_frame$CURRENT_AGE_DEID
plot(age_numeric, mart_resid,
     main = "Linearity: Age",
     xlab = "Age", ylab = "Martingale Residuals")
abline(h = 0, lty = 2)
lines(lowess(age_numeric, mart_resid), col = "red", lwd = 2)

msi_numeric <- model_frame$logMSI
plot(msi_numeric, mart_resid,
     main = "Linearity: MSI Score",
     xlab = "MSI Score", ylab = "Martingale Residuals")
abline(h = 0, lty = 2)
lines(lowess(msi_numeric, mart_resid), col = "red", lwd = 2)

tmb_numeric <- model_frame$logTMB
plot(tmb_numeric, mart_resid,
     main = "Linearity: TMB",
     xlab = "TMB Non-Synonymous", ylab = "Martingale Residuals")
abline(h = 0, lty = 2)
lines(lowess(tmb_numeric, mart_resid), col = "red", lwd = 2)

par(mfrow = c(1, 1))
```

```{r Outlier Analysis}

# deviance (standardized resids)
deviance_residuals <- residuals(model_adjusted, type = "deviance")

# flag any residuals greater than abs(3)
outliers <- which(abs(deviance_residuals) > 3)
n_outliers <- length(outliers)

cat("Number of outliers (|deviance residual| > 3):", n_outliers, "\n")
cat("Percentage of cohort:", round((n_outliers / nrow(survDF)) * 100, 1), "%\n")

if (n_outliers > 0 && n_outliers <= 10) {
  cat("\nOutlier patient IDs:\n")
  print(survDF$PATIENT_ID[outliers])
}

# Plot residuals
is_outlier <- abs(deviance_residuals) > 3
colors <- ifelse(is_outlier, "red", "black")

plot(fitted(model_adjusted), deviance_residuals,
     main = "Deviance Residuals vs Fitted Values",
     xlab = "Fitted Values (log hazard)",
     ylab = "Deviance Residuals",
     col = colors)
abline(h = 0, lty = 2)
abline(h = c(-3, 3), lty = 3, col = "red")

legend("topright", 
       legend = c("Normal", "Outlier"),
       col = c("black", "red"),
       pch = 1)
```

```{r Multicollinearity Check}

# make corr matrix (only numeric vars)
numeric_vars <- survDF |>
  dplyr::select(
    CURRENT_AGE_DEID, ageSquared,
    logMSI, logMSISquared,
    logTMB, logTMBSquared,
    ends_with("_mut"),
    OS_time, OS_event
  ) |>
  drop_na()

corr_matrix <- cor(numeric_vars)
print(round(corr_matrix, 3))

# Convert to long format for ggplot
corr_long <- melt(corr_matrix) |>
  filter(Var1 != Var2) |>  
  filter(abs(value) > 0.7)

# Create heatmap
ggplot(corr_long, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile(color = "black", size = 0.5) +
  geom_text(aes(label = round(value, 2)), color = "black", size = 3) +
  scale_fill_gradient2(low = "lightblue", mid = "white", high = "red",
                       limits = c(-1, 1), midpoint = 0) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        axis.text.y = element_text(size = 10),
        axis.title = element_blank()) +
  labs(title = "Correlation Matrix Heatmap (selected correlations > 0.7)",
       fill = "Correlation")
```

```{r Bootstrap CIs}

bootstrap_one_gene <- function(data, gene_name, indices) {
  boot_data <- data[indices, ]
  geneVar <- paste0(gene_name, "_mut")
  
  model_boot <- coxph(
    as.formula(paste(
      "Surv(OS_time, OS_event) ~", geneVar,
      "+ moa_consolidated +", geneVar, ":moa_consolidated",
      "+ CURRENT_AGE_DEID + ageSquared + GENDER + RACE + STAGE_HIGHEST_RECORDED",
      "+ logMSI + logMSISquared + logTMB + logTMBSquared"
    )),
    data = boot_data
  )
  
  coef_boot <- coef(model_boot)
  ix_pos <- grep(":moa_consolidated", names(coef_boot))
  return(exp(coef_boot[ix_pos]))
}

set.seed(107)

boot_ci_all_genes <- tibble()

for (gene in c("BRAF", "APC", "PIK3CA", "SMAD4")) { 
  
  cat("Bootstrapping:", gene, "...\n")
  
  boot_res <- boot(survDF, bootstrap_one_gene, R = 250, gene_name = gene)
  
  for (i in 1:length(boot_res$t0)) {
    interaction_name <- names(boot_res$t0)[i]
    
    ci_lower <- quantile(boot_res$t[, i], 0.025, na.rm = TRUE)
    ci_upper <- quantile(boot_res$t[, i], 0.975, na.rm = TRUE)
    point_est <- boot_res$t0[i]
    n_na <- sum(is.na(boot_res$t[, i]))
    
    boot_ci_all_genes <- bind_rows(
      boot_ci_all_genes,
      tibble(
        Gene = gene,
        Interaction = interaction_name,
        `Point HR` = round(point_est, 3),
        `CI Lower` = round(ci_lower, 3),
        `CI Upper` = round(ci_upper, 3),
        `95% CI` = paste0(round(ci_lower, 3), "-", round(ci_upper, 3)),
        `Failed Boots` = n_na,
        Robust = ifelse(ci_lower > 1 | ci_upper < 1, "YES", "NO")
      )
    )
  }
}

print(boot_ci_all_genes)
```

```{r Sensitivity Analysis- Removing Outliers}

deviance_resid <- residuals(model_adjusted, type = "deviance")
outlier_indices <- which(abs(deviance_resid) > 3)
n_outliers <- length(outlier_indices)
pct_outliers <- (n_outliers / nrow(survDF)) * 100

cat("Number of outliers:", n_outliers, "(", round(pct_outliers, 2), "%)\n")

if (n_outliers > 0 & n_outliers <= 20) {
  cat("Outlier patient IDs:\n")
  print(survDF$PATIENT_ID[outlier_indices])
  cat("\n")
}

# Create dataset without outliers
survDF_no_outliers <- survDF[-outlier_indices, ]

cat("Original sample size:", nrow(survDF), "\n")
cat("Sample size without outliers:", nrow(survDF_no_outliers), "\n")
cat("Patients removed:", nrow(survDF) - nrow(survDF_no_outliers), "\n\n")

# Refit key gene models without outliers
sensitivity_comparison <- tibble()

for (gene in c("BRAF", "APC", "PIK3CA", "SMAD4")) {
  
  geneVar <- paste0(gene, "_mut")
  
  model_no_out <- coxph(
    as.formula(paste(
      "Surv(OS_time, OS_event) ~", geneVar,
      "+ moa_consolidated +", geneVar, ":moa_consolidated",
      "+ CURRENT_AGE_DEID + ageSquared + GENDER + RACE + STAGE_HIGHEST_RECORDED",
      "+ logMSI + logMSISquared + logTMB + logTMBSquared"
    )),
    data = survDF_no_outliers
  )
  
  orig_coef <- coef(uniRes_improved[[gene]]$model)
  no_out_coef <- coef(model_no_out)
  
  orig_ix <- grep(":moa_consolidated", names(orig_coef))
  no_out_ix <- grep(":moa_consolidated", names(no_out_coef))
  
  for (j in 1:length(orig_ix)) {
    interaction_name <- names(orig_coef)[orig_ix[j]]
    
    orig_hr <- exp(orig_coef[orig_ix[j]])
    no_out_hr <- exp(no_out_coef[no_out_ix[j]])
    pct_change <- ((no_out_hr - orig_hr) / orig_hr) * 100
    
    sensitivity_comparison <- bind_rows(
      sensitivity_comparison,
      tibble(
        Gene = gene,
        Interaction = str_replace_all(interaction_name, "_mut:moa_consolidated", "") |>
          str_replace("Chemo", "Chemo ") |> str_trim(),
        `Original HR` = round(orig_hr, 3),
        `No Outliers HR` = round(no_out_hr, 3),
        `% Change` = round(pct_change, 1),
        Stable = ifelse(abs(pct_change) < 10, "YES", "NO")
      )
    )
  }
}

print(sensitivity_comparison)
```

---

# ADDITIONAL ANALYSES: Addressing PH Violations and Visualization

## Kaplan-Meier Curves: BRAF and PIK3CA

```{r Kaplan-Meier BRAF}
#| fig-width: 14
#| fig-height: 6

# BRAF - stratified by mutation and treatment
survDF_braf <- survDF |>
  mutate(
    BRAF_Status = ifelse(BRAF_mut == 1, "Mutant", "Wildtype"),
    Group = paste0("BRAF ", BRAF_Status, "\n", moa_consolidated)
  )

fit_braf <- survfit(Surv(OS_time, OS_event) ~ Group, data = survDF_braf)

p_braf <- ggsurvplot(
  fit_braf,
  data = survDF_braf,
  risk.table = TRUE,
  pval = TRUE,
  conf.int = TRUE,
  title = "Kaplan-Meier: Overall Survival by BRAF Mutation Status and Treatment",
  xlab = "Months from Treatment Start",
  ylab = "Overall Survival Probability",
  legend.title = "Group",
  ggtheme = theme_minimal()
)

print(p_braf)
```

```{r Kaplan-Meier PIK3CA}
#| fig-width: 14
#| fig-height: 6

# PIK3CA - stratified by mutation and treatment
survDF_pik3ca <- survDF |>
  mutate(
    PIK3CA_Status = ifelse(PIK3CA_mut == 1, "Mutant", "Wildtype"),
    Group = paste0("PIK3CA ", PIK3CA_Status, "\n", moa_consolidated)
  )

fit_pik3ca <- survfit(Surv(OS_time, OS_event) ~ Group, data = survDF_pik3ca)

p_pik3ca <- ggsurvplot(
  fit_pik3ca,
  data = survDF_pik3ca,
  risk.table = TRUE,
  pval = TRUE,
  conf.int = TRUE,
  title = "Kaplan-Meier: Overall Survival by PIK3CA Mutation Status and Treatment",
  xlab = "Months from Treatment Start",
  ylab = "Overall Survival Probability",
  legend.title = "Group",
  ggtheme = theme_minimal()
)

print(p_pik3ca)
```

---

## Restricted Mean Survival Time (RMST) Analysis


## Restricted Mean Survival Time (RMST) Analysis
```{r RMST Analysis - requires survRM2}
#| warning: false
#| message: false

library(survRM2)

cat("\n=== RESTRICTED MEAN SURVIVAL TIME (RMST) ANALYSIS ===\n")
cat("RMST measures average survival over a restricted time period.\n")
cat("Does NOT assume proportional hazards.\n\n")

rmst_results <- tibble()

# Compare BRAF mutant vs WT in each treatment group
for (moa_group in unique(survDF$moa_consolidated)) {
  
  df_subset <- survDF |> 
    filter(moa_consolidated == moa_group) |>
    filter(!is.na(BRAF_mut))
  
  if (nrow(df_subset) < 10) next
  
  # Calculate tau separately for each subset (min of max follow-up in each arm)
  max_time_mut <- max(df_subset$OS_time[df_subset$BRAF_mut == 1], na.rm = TRUE)
  max_time_wt <- max(df_subset$OS_time[df_subset$BRAF_mut == 0], na.rm = TRUE)
  tau_subset <- min(max_time_mut, max_time_wt) * 0.9  # Use 90% of minimum max time
  
  if (tau_subset <= 0) next
  
  cat("===== BRAF in", moa_group, "=====\n")
  cat("Tau (restricted time):", round(tau_subset, 1), "months\n")
  
  rmst_braf <- rmst2(
    time = df_subset$OS_time,
    status = df_subset$OS_event,
    arm = df_subset$BRAF_mut,
    tau = tau_subset
  )
  
  print(rmst_braf)
  cat("\n")
  
  rmst_results <- bind_rows(
    rmst_results,
    tibble(
      Gene = "BRAF",
      Treatment = moa_group,
      Tau_Months = round(tau_subset, 1),
      RMST_Diff = rmst_braf$unadjusted.result[1, 1],
      RMST_SE = rmst_braf$unadjusted.result[1, 2],
      P_value = rmst_braf$unadjusted.result[1, 4]
    )
  )
}

cat("\n")

# Compare PIK3CA mutant vs WT in each treatment group
for (moa_group in unique(survDF$moa_consolidated)) {
  
  df_subset <- survDF |> 
    filter(moa_consolidated == moa_group) |>
    filter(!is.na(PIK3CA_mut))
  
  if (nrow(df_subset) < 10) next
  
  # Calculate tau separately for each subset
  max_time_mut <- max(df_subset$OS_time[df_subset$PIK3CA_mut == 1], na.rm = TRUE)
  max_time_wt <- max(df_subset$OS_time[df_subset$PIK3CA_mut == 0], na.rm = TRUE)
  tau_subset <- min(max_time_mut, max_time_wt) * 0.9
  
  if (tau_subset <= 0) next
  
  cat("===== PIK3CA in", moa_group, "=====\n")
  cat("Tau (restricted time):", round(tau_subset, 1), "months\n")
  
  rmst_pik3ca <- rmst2(
    time = df_subset$OS_time,
    status = df_subset$OS_event,
    arm = df_subset$PIK3CA_mut,
    tau = tau_subset
  )
  
  print(rmst_pik3ca)
  cat("\n")
  
  rmst_results <- bind_rows(
    rmst_results,
    tibble(
      Gene = "PIK3CA",
      Treatment = moa_group,
      Tau_Months = round(tau_subset, 1),
      RMST_Diff = rmst_pik3ca$unadjusted.result[1, 1],
      RMST_SE = rmst_pik3ca$unadjusted.result[1, 2],
      P_value = rmst_pik3ca$unadjusted.result[1, 4]
    )
  )
}

# Display results table
if (nrow(rmst_results) > 0) {
  cat("\n=== RMST SUMMARY TABLE ===\n")
  rmst_summary <- rmst_results |>
    mutate(
      `RMST Diff (months)` = round(RMST_Diff, 2),
      `95% CI` = paste0("±", round(1.96 * RMST_SE, 2)),
      `P-value` = round(P_value, 4)
    ) |>
    select(Gene, Treatment, Tau_Months, `RMST Diff (months)`, `95% CI`, `P-value`)
  
  print(rmst_summary)
  
  rmst_summary |>
    gt() |>
    tab_header(
      title = "Table: Restricted Mean Survival Time Analysis",
      subtitle = "Restricted to 90% of minimum max follow-up per group"
    ) |>
    tab_source_note("RMST Diff = difference in mean survival time (Mutant - Wildtype). Positive = mutants survive longer.")
} else {
  cat("No RMST results available - insufficient sample size in groups.\n")
}

```

---

**Key changes:**

✓ `tau` calculated **separately for each gene-treatment combination**
✓ Uses 90% of the **minimum max follow-up** between the two arms
✓ Skips groups with insufficient data
✓ Reports `Tau_Months` so you can see what was used

This should resolve the error!

---

## Interaction Plots: Visualizing Gene-Treatment Patterns

## Interaction Plots: Visualizing Gene-Treatment Patterns
```{r Interaction Plots - BRAF and PIK3CA}
#| fig-width: 12
#| fig-height: 6

# Filter stratified results for top genes (remove _mut since it's not in the data)
interaction_plot_data <- stratHR |>
  filter(Gene %in% c("BRAF", "PIK3CA"))

# Only proceed if we have data
if (nrow(interaction_plot_data) > 0) {
  
  # Plot interaction
  p_interact <- interaction_plot_data |>
    ggplot(aes(x = Treatment, y = log(HR), color = Gene, group = Gene)) +
    geom_point(size = 4) +
    geom_line(size = 1.2) +
    geom_errorbar(aes(ymin = log(CI_lower), ymax = log(CI_upper)), 
                  width = 0.15, size = 1) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "gray50", size = 0.8) +
    facet_wrap(~Gene, nrow = 1) +
    scale_color_manual(values = c("BRAF" = "#d62728", "PIK3CA" = "#2ca02c")) +
    labs(
      title = "Interaction Patterns: Mutation Effects Vary by Treatment",
      x = "Treatment Mechanism of Action",
      y = "log(Hazard Ratio)",
      color = "Gene",
      subtitle = "Non-parallel lines indicate gene-treatment interaction"
    ) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1, size = 11),
      plot.title = element_text(face = "bold", size = 13),
      legend.position = "bottom"
    )
  
  print(p_interact)
  
} else {
  cat("No data available for interaction plots\n")
}
```


## AFT (Accelerated Failure Time) Models - Robustness Check
```{r AFT Models - Weibull}
#| warning: false

cat("\n=== ACCELERATED FAILURE TIME (AFT) MODELS ===\n")
cat("Parametric alternative to Cox PH models.\n")
cat("More robust when PH assumption violated.\n\n")

# Fit Weibull AFT models for BRAF and PIK3CA
aft_results <- tibble()

for (gene in c("BRAF", "PIK3CA")) {
  geneVar <- paste0(gene, "_mut")
  
  cat("Fitting AFT model for", gene, "...\n")
  
  aft_model <- survreg(
    as.formula(paste(
      "Surv(OS_time, OS_event) ~", geneVar, 
      "+ moa_consolidated +", geneVar, ":moa_consolidated",
      "+ CURRENT_AGE_DEID + ageSquared + GENDER + RACE + STAGE_HIGHEST_RECORDED",
      "+ logMSI + logTMB + logMSISquared + logTMBSquared"
    )),
    data = survDF,
    dist = "weibull"
  )
  
  aft_summary <- summary(aft_model)
  
  # Check column names
  cat("Column names in aft_summary$table:\n")
  print(colnames(aft_summary$table))
  cat("\n")
  
  # Extract interaction coefficients - extract by column position instead
  int_coefs_df <- as.data.frame(aft_summary$table) |>
    rownames_to_column("Variable") |>
    filter(str_detect(Variable, ":moa_consolidated"))
  
  if (nrow(int_coefs_df) > 0) {
    # Extract columns by position: Value (1), Std. Error (2), z-value (3), p-value (4)
    int_coefs <- int_coefs_df |>
      mutate(
        Gene = gene,
        AFT_Coef = int_coefs_df[[2]],  # Value column
        SE = int_coefs_df[[3]],  # Std. Error column
        P_value = int_coefs_df[[5]],  # p-value column (usually 5th)
        HR_equiv = exp(-AFT_Coef)
      ) |>
      select(Gene, Variable, AFT_Coef, HR_equiv, P_value)
    
    aft_results <- bind_rows(aft_results, int_coefs)
    
    # Compare to Cox
    cox_model <- uniRes_improved[[gene]]$model
    cox_p <- summary(cox_model)$coefficients[grep(":moa", rownames(summary(cox_model)$coefficients)), "Pr(>|z|)"]
    
    cat("Gene:", gene, "\n")
    cat("  Cox interaction p-values:", round(cox_p, 4), "\n")
    cat("  AFT interaction p-values:", round(int_coefs$P_value, 4), "\n\n")
  }
}

# Display comparison table
cat("\n=== AFT vs COX COMPARISON ===\n")
if (nrow(aft_results) > 0) {
  aft_results |>
    mutate(
      HR_equiv = round(HR_equiv, 3),
      P_value = round(P_value, 4)
    ) |>
    select(Gene, Variable, HR_equiv, P_value) |>
    arrange(Gene) |>
    gt() |>
    tab_header(
      title = "Comparison: AFT vs Cox Models",
      subtitle = "Robustness check for PH assumption violations"
    ) |>
    tab_source_note("HR_equiv = Weibull AFT coefficient converted to hazard ratio scale for comparison.")
} else {
  cat("No AFT results available.\n")
}
```


## Subgroup Forest Plot: All Genes × Treatments

```{r Subgroup Forest Plot}
#| fig-width: 13
#| fig-height: 9

# Create forest plot data
forest_data <- stratHR |>
  mutate(
    Gene = str_remove(Gene, "_mut"),
    Label = paste0(Gene, " (", Treatment, ")")
  ) |>
  arrange(HR) |>
  mutate(Label = fct_inorder(Label))

# Plot
p_forest <- forest_data |>
  ggplot(aes(y = Label, x = HR)) +
  geom_point(size = 3, color = "#1f77b4", alpha = 0.7) +
  geom_errorbarh(aes(xmin = CI_lower, xmax = CI_upper), 
                 height = 0.2, size = 0.8, color = "#1f77b4") +
  geom_vline(xintercept = 1, linetype = "dashed", color = "red", size = 1) +
  scale_x_log10(breaks = c(0.5, 1, 2, 4)) +
  labs(
    title = "Subgroup Analysis: Mutation Effects by Treatment Group",
    x = "Hazard Ratio (log scale, 95% CI)",
    y = "Gene × Treatment Combination",
    subtitle = "Positive: worse survival with mutation; <1: better survival with mutation"
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 9),
    plot.title = element_text(face = "bold", size = 13),
    panel.grid.major.x = element_line(color = "lightgray", size = 0.3)
  )

print(p_forest)
```



```{r report table and figure making}

# ============================================================
# TABLE 1
# ============================================================
survDF |>
  mutate(
    moa_consolidated = factor(moa_consolidated, 
                               levels = c("Chemo", "Chemo+Targeted", "Other/IO")),
    GENDER = factor(GENDER),
    RACE = factor(RACE),
    STAGE_HIGHEST_RECORDED = factor(STAGE_HIGHEST_RECORDED)
  ) |>
  dplyr::select(
    moa_consolidated,
    CURRENT_AGE_DEID, GENDER, RACE, STAGE_HIGHEST_RECORDED,
    MSI_SCORE, TMB_NONSYNONYMOUS,
    TP53_mut, APC_mut, KRAS_mut, SMAD4_mut, PIK3CA_mut,
    TCF7L2_mut, FBXW7_mut, SOX9_mut, BRAF_mut, ARID1A_mut
  ) |>
  tbl_summary(
    by = moa_consolidated,
    label = list(
      CURRENT_AGE_DEID ~ "Age at Diagnosis",
      GENDER ~ "Sex",
      RACE ~ "Race",
      STAGE_HIGHEST_RECORDED ~ "Highest Stage",
      MSI_SCORE ~ "MSI Score",
      TMB_NONSYNONYMOUS ~ "TMB (Non-Synonymous)",
      TP53_mut ~ "TP53 Mutant, n (%)",
      APC_mut ~ "APC Mutant, n (%)",
      KRAS_mut ~ "KRAS Mutant, n (%)",
      SMAD4_mut ~ "SMAD4 Mutant, n (%)",
      PIK3CA_mut ~ "PIK3CA Mutant, n (%)",
      TCF7L2_mut ~ "TCF7L2 Mutant, n (%)",
      FBXW7_mut ~ "FBXW7 Mutant, n (%)",
      SOX9_mut ~ "SOX9 Mutant, n (%)",
      BRAF_mut ~ "BRAF Mutant, n (%)",
      ARID1A_mut ~ "ARID1A Mutant, n (%)"
    ),
    statistic = list(
      all_continuous() ~ "{median} ({p25}, {p75})",
      all_categorical() ~ "{n} ({p}%)"
    ),
    missing = "no"
  ) |>
  add_overall() |>
  add_p() |>
  bold_labels() |>
  modify_caption("**Table 1. Baseline Characteristics by First-Line Treatment Group**") |>
  as_gt() |>
  tab_options(
    table.font.size = px(9),
    table.width = pct(100),
    column_labels.font.size = px(9),
    heading.title.font.size = px(11),
    heading.subtitle.font.size = px(9),
    data_row.padding = px(2)
  )

# ============================================================
# TABLE 2
# ============================================================
chemo_results <- stratHR |>
  filter(Treatment == "Chemo") |>
  mutate(
    `HR (95% CI)` = paste0(
      round(HR, 2), " (",
      round(CI_lower, 2), "\u2013",
      round(CI_upper, 2), ")"
    ),
    `P-value` = case_when(
      P_value < 0.001 ~ "<0.001",
      TRUE ~ as.character(round(P_value, 3))
    ),
    Significant = ifelse(P_value < 0.05, "*", "")
  ) |>
  arrange(P_value) |>
  dplyr::select(Gene, `HR (95% CI)`, `P-value`, Significant, P_value)

chemo_results |>
  dplyr::select(Gene, `HR (95% CI)`, `P-value`, Significant) |>
  gt() |>
  tab_header(
    title = "Table 2. Cox PH Results: Chemotherapy Group",
    subtitle = "Adjusted for age, sex, race, stage, logMSI, logTMB"
  ) |>
  cols_label(
    Gene = "Gene",
    `HR (95% CI)` = "HR (95% CI)",
    `P-value` = "P-value",
    Significant = ""
  ) |>
  tab_footnote(
    footnote = "HR > 1 = worse OS in mutation carriers. * p < 0.05.",
    locations = cells_column_labels(columns = `HR (95% CI)`)
  ) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_body(rows = chemo_results$P_value < 0.05)
  ) |>
  tab_options(
    table.font.size = px(10),
    table.width = pct(70),
    heading.title.font.size = px(11),
    heading.subtitle.font.size = px(9),
    data_row.padding = px(3)
  ) |>
  opt_stylize(style = 2)

# ============================================================
# FIGURE 1 — Forest Plot of Gene x MoA Interaction HRs
# ============================================================
forest_ix <- ixTab |>
  mutate(
    Treatment = case_when(
      str_detect(Interaction_Term, "Chemo\\+Targeted") ~ "Chemo + Targeted",
      str_detect(Interaction_Term, "Other") ~ "Other/IO",
      TRUE ~ "Other"
    ),
    Label = paste0(Gene, " \u00d7 ", Treatment),
    FDR_label = paste0("p=", round(P_value, 3),
                        "\n(FDR p=", round(P_value_fdr, 3), ")")
  ) |>
  arrange(HR) |>
  mutate(Label = fct_inorder(Label))

ggplot(forest_ix, aes(y = Label, x = HR, color = Treatment)) +
  geom_point(size = 3) +
  geom_errorbar(
    aes(xmin = CI_lower, xmax = CI_upper),
    width = 0.25, linewidth = 0.8,
    orientation = "y"
  ) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "gray40", linewidth = 0.9) +
  geom_text(
    aes(label = FDR_label, x = pmax(CI_upper, HR) + 0.05),
    hjust = 0, size = 2.8, color = "gray30"
  ) +
  scale_x_log10(
    breaks = c(0.25, 0.5, 1, 2, 4, 8),
    labels = c("0.25", "0.5", "1", "2", "4", "8")
  ) +
  scale_color_manual(values = c("Chemo + Targeted" = "#1f77b4", "Other/IO" = "#ff7f0e")) +
  labs(
    title = "Figure 1. Forest Plot of Gene \u00d7 Treatment Interaction Hazard Ratios",
    subtitle = "Reference group: Cytotoxic Chemotherapy. Dashed line = null (HR = 1).",
    x = "Interaction Hazard Ratio (log scale, 95% CI)",
    y = "Gene \u00d7 Treatment Interaction",
    color = "Treatment Group",
    caption = "FDR correction applied using Benjamini-Hochberg method across 20 tests."
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom",
    plot.title = element_text(face = "bold"),
    axis.text.y = element_text(size = 10),
    plot.caption = element_text(size = 9, color = "gray40")
  ) +
  coord_cartesian(clip = "off") +
  theme(plot.margin = margin(r = 100))

# ============================================================
# FIGURE 2 — KM Curves: BRAF and APC in Chemotherapy Group
# ============================================================
survDF_chemo <- survDF |>
  filter(moa_consolidated == "Chemo") |>
  mutate(
    BRAF_label = ifelse(BRAF_mut == 1, "BRAF Mutant", "BRAF Wildtype"),
    APC_label  = ifelse(APC_mut == 1,  "APC Mutant",  "APC Wildtype")
  )

fit_braf_chemo <- survfit(Surv(OS_time, OS_event) ~ BRAF_label, data = survDF_chemo)

p_braf_chemo <- ggsurvplot(
  fit_braf_chemo,
  data = survDF_chemo,
  risk.table = TRUE,
  pval = TRUE,
  conf.int = TRUE,
  palette = c("#d62728", "#aec7e8"),
  title = "BRAF Mutation Status \u2014 Chemotherapy Group",
  xlab = "Months from First-Line Treatment Start",
  ylab = "Overall Survival Probability",
  legend.title = "",
  legend.labs = c("BRAF Mutant", "BRAF Wildtype"),
  ggtheme = theme_minimal(base_size = 12),
  risk.table.height = 0.28,
  tables.theme = theme_cleantable()
)

fit_apc_chemo <- survfit(Surv(OS_time, OS_event) ~ APC_label, data = survDF_chemo)

p_apc_chemo <- ggsurvplot(
  fit_apc_chemo,
  data = survDF_chemo,
  risk.table = TRUE,
  pval = TRUE,
  conf.int = TRUE,
  palette = c("#2ca02c", "#98df8a"),
  title = "APC Mutation Status \u2014 Chemotherapy Group",
  xlab = "Months from First-Line Treatment Start",
  ylab = "Overall Survival Probability",
  legend.title = "",
  legend.labs = c("APC Mutant", "APC Wildtype"),
  ggtheme = theme_minimal(base_size = 12),
  risk.table.height = 0.28,
  tables.theme = theme_cleantable()
)

arrange_ggsurvplots(
  list(p_braf_chemo, p_apc_chemo),
  ncol = 2,
  title = "Figure 2. Kaplan-Meier Curves by Mutation Status in the Chemotherapy Group"
)

# ============================================================
# FIGURE 3 — Landmark Analysis + Bootstrap for PIK3CA
# ============================================================
pik3ca_landmark <- landmark_results |>
  filter(Gene == "PIK3CA") |>
  mutate(
    Treatment = case_when(
      str_detect(Variable, "Chemo\\+Targeted") ~ "Chemo + Targeted",
      str_detect(Variable, "Other")            ~ "Other/IO",
      TRUE ~ "Unknown"
    ),
    Period = factor(Period, levels = c("Early", "Late")),
    sig_label = case_when(
      P_value_fdr < 0.05 ~ paste0("FDR p=", round(P_value_fdr, 3), " *"),
      P_value < 0.05     ~ paste0("p=", round(P_value, 3), " (FDR p=", round(P_value_fdr, 3), ")"),
      TRUE               ~ paste0("p=", round(P_value, 3))
    )
  )

p_landmark <- ggplot(pik3ca_landmark,
                     aes(x = Period, y = HR, color = Treatment, group = Treatment)) +
  geom_point(size = 4) +
  geom_line(linewidth = 1.1, linetype = "dashed") +
  geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper), width = 0.12, linewidth = 0.9) +
  geom_text(aes(label = sig_label), vjust = -1.2, size = 3.2) +
  geom_hline(yintercept = 1, linetype = "dotted", color = "gray50") +
  scale_color_manual(values = c("Chemo + Targeted" = "#1f77b4", "Other/IO" = "#ff7f0e")) +
  scale_y_log10() +
  labs(
    title = "A. PIK3CA \u00d7 Treatment \u2014 Landmark Analysis",
    x = "Follow-Up Period (Split at Median: 35.5 Months)",
    y = "Interaction HR (log scale, 95% CI)",
    color = "Treatment"
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"), legend.position = "bottom")

pik3ca_boot <- boot_ci_all_genes |>
  filter(Gene == "PIK3CA") |>
  mutate(
    Treatment = case_when(
      str_detect(Interaction, "Chemo.Targeted") ~ "Chemo + Targeted",
      str_detect(Interaction, "Other")          ~ "Other/IO",
      TRUE ~ Interaction
    )
  )

p_boot <- ggplot(pik3ca_boot, aes(y = Treatment, x = `Point HR`, color = Robust)) +
  geom_point(size = 4) +
  geom_errorbar(
    aes(xmin = `CI Lower`, xmax = `CI Upper`),
    width = 0.2, linewidth = 0.9,
    orientation = "y"
  ) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "gray40", linewidth = 0.9) +
  scale_color_manual(values = c("YES" = "#d62728", "NO" = "#aec7e8"),
                     labels = c("YES" = "Robust (CI excludes 1)", "NO" = "Not robust")) +
  labs(
    title = "B. PIK3CA \u2014 Bootstrap Validation (250 Resamples)",
    x = "Interaction HR (95% Bootstrapped CI)",
    y = "Treatment Comparison",
    color = "Robustness"
  ) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(face = "bold"), legend.position = "bottom")

library(patchwork)
p_landmark + p_boot +
  plot_annotation(
    title = "Figure 3. PIK3CA \u00d7 Chemo+Targeted: Landmark Analysis and Bootstrap Validation",
    caption = "Left: Interaction HRs by landmark period; * = FDR p < 0.05. Right: Bootstrap 95% CIs from 250 resamples.",
    theme = theme(plot.title = element_text(face = "bold", size = 13))
  )

# ============================================================
# TABLE 3 — RMST (add APC first, then display)
# ============================================================
for (moa_group in unique(survDF$moa_consolidated)) {
  df_subset <- survDF |> filter(moa_consolidated == moa_group, !is.na(APC_mut))
  if (nrow(df_subset) < 10) next
  tau_apc <- min(
    max(df_subset$OS_time[df_subset$APC_mut == 1], na.rm = TRUE),
    max(df_subset$OS_time[df_subset$APC_mut == 0], na.rm = TRUE)
  ) * 0.9
  if (tau_apc <= 0) next
  r <- rmst2(df_subset$OS_time, df_subset$OS_event, df_subset$APC_mut, tau = tau_apc)
  rmst_results <- bind_rows(rmst_results, tibble(
    Gene = "APC", Treatment = moa_group,
    Tau_Months = round(tau_apc, 1),
    RMST_Diff = r$unadjusted.result[1, 1],
    RMST_SE   = r$unadjusted.result[1, 2],
    P_value   = r$unadjusted.result[1, 4]
  ))
}

rmst_results |>
  filter(Gene %in% c("APC", "BRAF", "PIK3CA")) |>
  mutate(
    `RMST Diff (mo.)` = round(RMST_Diff, 2),
    `95% CI` = paste0(
      round(RMST_Diff - 1.96 * RMST_SE, 2), " to ",
      round(RMST_Diff + 1.96 * RMST_SE, 2)
    ),
    `P-value` = case_when(P_value < 0.001 ~ "<0.001", TRUE ~ as.character(round(P_value, 3))),
    Gene = factor(Gene, levels = c("BRAF", "PIK3CA", "APC"))
  ) |>
  arrange(Gene, Treatment) |>
  dplyr::select(Gene, Treatment, Tau_Months, `RMST Diff (mo.)`, `95% CI`, `P-value`) |>
  gt() |>
  tab_header(
    title = "Table 3. RMST Analysis",
    subtitle = "Mutant vs. Wildtype by treatment group"
  ) |>
  cols_label(
    Tau_Months = "Tau (mo.)",
    `RMST Diff (mo.)` = "RMST Diff (mo.)",
    `95% CI` = "95% CI",
    `P-value` = "P-value"
  ) |>
  tab_footnote(
    footnote = "Diff = Mutant \u2212 Wildtype. Negative = worse OS in mutation carriers. Tau = 90% of minimum arm max follow-up.",
    locations = cells_column_labels(columns = `RMST Diff (mo.)`)
  ) |>
  tab_row_group(label = "BRAF",   rows = Gene == "BRAF")   |>
  tab_row_group(label = "PIK3CA", rows = Gene == "PIK3CA") |>
  tab_row_group(label = "APC",    rows = Gene == "APC")    |>
  tab_options(
    table.font.size = px(9),
    table.width = pct(95),
    heading.title.font.size = px(11),
    heading.subtitle.font.size = px(9),
    column_labels.font.size = px(9),
    data_row.padding = px(2),
    row_group.font.weight = "bold"
  ) |>
  opt_stylize(style = 2)
  
```


```{r figures try again}

# ============================================================
# FIGURE 1 — Forest Plot of Gene x MoA Interaction HRs
# ============================================================
#| fig-width: 14
#| fig-height: 10

library(ggrepel)

forest_ix <- ixTab |>
  mutate(
    Treatment = case_when(
      str_detect(Interaction_Term, "Chemo\\+Targeted") ~ "Chemo + Targeted",
      str_detect(Interaction_Term, "Other") ~ "Other/IO",
      TRUE ~ "Other"
    ),
    Label = paste0(Gene, " \u00d7 ", Treatment),
    FDR_label = paste0("p=", round(P_value, 3),
                       "  FDR p=", round(P_value_fdr, 3))
  ) |>
  arrange(HR) |>
  mutate(Label = fct_inorder(Label))

ggplot(forest_ix, aes(y = Label, x = HR, color = Treatment)) +
  geom_point(size = 3) +
  geom_errorbar(
    aes(xmin = CI_lower, xmax = CI_upper),
    width = 0.4,
    linewidth = 0.8,
    orientation = "y"
  ) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "gray40", linewidth = 0.9) +
  geom_text(
    aes(label = FDR_label, x = CI_upper + 0.1),
    hjust = 0,
    size = 2.6,
    color = "gray30"
  ) +
  scale_x_log10(
    breaks = c(0.25, 0.5, 1, 2, 4, 8),
    labels = c("0.25", "0.5", "1", "2", "4", "8"),
    limits = c(0.1, 20)
  ) +
  scale_color_manual(values = c("Chemo + Targeted" = "#1f77b4", "Other/IO" = "#ff7f0e")) +
  labs(
    title = "Figure 1. Forest Plot of Gene \u00d7 Treatment Interaction Hazard Ratios",
    subtitle = "Reference group: Cytotoxic Chemotherapy. Dashed line = null (HR = 1).",
    x = "Interaction Hazard Ratio (log scale, 95% CI)",
    y = "Gene \u00d7 Treatment Interaction",
    color = "Treatment Group",
    caption = "FDR correction applied using Benjamini-Hochberg method across 20 tests."
  ) +
  theme_minimal(base_size = 11) +
  theme(
    legend.position = "bottom",
    plot.title = element_text(face = "bold", size = 12),
    axis.text.y = element_text(size = 9),
    axis.text.x = element_text(size = 9),
    plot.caption = element_text(size = 8, color = "gray40"),
    plot.margin = margin(t = 10, r = 160, b = 10, l = 10),
    panel.grid.minor = element_blank()
  ) +
  coord_cartesian(clip = "off")

# ============================================================
# FIGURE 2 — KM Curves: BRAF and APC in Chemotherapy Group
# ============================================================
#| fig-width: 16
#| fig-height: 8

survDF_chemo <- survDF |>
  filter(moa_consolidated == "Chemo") |>
  mutate(
    BRAF_label = ifelse(BRAF_mut == 1, "BRAF Mutant", "BRAF Wildtype"),
    APC_label  = ifelse(APC_mut == 1,  "APC Mutant",  "APC Wildtype")
  )

fit_braf_chemo <- survfit(Surv(OS_time, OS_event) ~ BRAF_label, data = survDF_chemo)

p_braf_chemo <- ggsurvplot(
  fit_braf_chemo,
  data = survDF_chemo,
  risk.table = TRUE,
  pval = TRUE,
  pval.size = 4,
  conf.int = TRUE,
  palette = c("#d62728", "#aec7e8"),
  title = "BRAF Mutation \u2014 Chemotherapy",
  xlab = "Months from First-Line Treatment Start",
  ylab = "Overall Survival Probability",
  legend.title = "",
  legend.labs = c("BRAF Mutant", "BRAF Wildtype"),
  ggtheme = theme_minimal(base_size = 11),
  font.main = c(11, "bold"),
  font.x = c(10, "plain"),
  font.y = c(10, "plain"),
  font.tickslab = c(9, "plain"),
  font.legend = c(9, "plain"),
  risk.table.height = 0.25,
  risk.table.fontsize = 3,
  tables.theme = theme_cleantable(),
  break.time.by = 50,
  xlim = c(0, 400)
)

fit_apc_chemo <- survfit(Surv(OS_time, OS_event) ~ APC_label, data = survDF_chemo)

p_apc_chemo <- ggsurvplot(
  fit_apc_chemo,
  data = survDF_chemo,
  risk.table = TRUE,
  pval = TRUE,
  pval.size = 4,
  conf.int = TRUE,
  palette = c("#2ca02c", "#98df8a"),
  title = "APC Mutation \u2014 Chemotherapy",
  xlab = "Months from First-Line Treatment Start",
  ylab = "Overall Survival Probability",
  legend.title = "",
  legend.labs = c("APC Mutant", "APC Wildtype"),
  ggtheme = theme_minimal(base_size = 11),
  font.main = c(11, "bold"),
  font.x = c(10, "plain"),
  font.y = c(10, "plain"),
  font.tickslab = c(9, "plain"),
  font.legend = c(9, "plain"),
  risk.table.height = 0.25,
  risk.table.fontsize = 3,
  tables.theme = theme_cleantable(),
  break.time.by = 50,
  xlim = c(0, 400)
)

arrange_ggsurvplots(
  list(p_braf_chemo, p_apc_chemo),
  ncol = 2,
  nrow = 1,
  title = "Figure 2. Kaplan-Meier Curves by Mutation Status in the Chemotherapy Group",
  print = TRUE
)

# ============================================================
# FIGURE 3 — Landmark Analysis + Bootstrap for PIK3CA
# ============================================================
#| fig-width: 14
#| fig-height: 6

pik3ca_landmark <- landmark_results |>
  filter(Gene == "PIK3CA") |>
  mutate(
    Treatment = case_when(
      str_detect(Variable, "Chemo\\+Targeted") ~ "Chemo + Targeted",
      str_detect(Variable, "Other")            ~ "Other/IO",
      TRUE ~ "Unknown"
    ),
    Period = factor(Period, levels = c("Early", "Late")),
    sig_label = case_when(
      P_value_fdr < 0.05 ~ paste0("FDR p=", round(P_value_fdr, 3), " *"),
      P_value < 0.05     ~ paste0("p=", round(P_value, 3),
                                   "\n(FDR p=", round(P_value_fdr, 3), ")"),
      TRUE               ~ paste0("p=", round(P_value, 3))
    )
  )

p_landmark <- ggplot(pik3ca_landmark,
                     aes(x = Period, y = HR, color = Treatment, group = Treatment)) +
  geom_point(size = 3.5) +
  geom_line(linewidth = 1.0, linetype = "dashed") +
  geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper),
                width = 0.10, linewidth = 0.8) +
  geom_text(
    aes(label = sig_label),
    vjust = -1.0,
    size = 3.0,
    lineheight = 0.9
  ) +
  geom_hline(yintercept = 1, linetype = "dotted", color = "gray50") +
  scale_color_manual(values = c("Chemo + Targeted" = "#1f77b4",
                                "Other/IO" = "#ff7f0e")) +
  scale_y_log10(
    breaks = c(0.1, 0.5, 1, 5, 10, 50, 100),
    labels = c("0.1", "0.5", "1", "5", "10", "50", "100"),
    limits = c(0.04, 200)
  ) +
  expand_limits(y = c(0.04, 200)) +
  labs(
    title = "A. PIK3CA \u00d7 Treatment \u2014 Landmark Analysis",
    x = "Follow-Up Period (Split at Median: 35.5 Months)",
    y = "Interaction HR (log scale, 95% CI)",
    color = "Treatment"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title = element_text(face = "bold", size = 11),
    legend.position = "bottom",
    legend.text = element_text(size = 9),
    axis.text = element_text(size = 9),
    plot.margin = margin(t = 20, r = 10, b = 10, l = 10)
  )

pik3ca_boot <- boot_ci_all_genes |>
  filter(Gene == "PIK3CA") |>
  mutate(
    Treatment = case_when(
      str_detect(Interaction, "Chemo.Targeted") ~ "Chemo + Targeted",
      str_detect(Interaction, "Other")          ~ "Other/IO",
      TRUE ~ Interaction
    )
  )

p_boot <- ggplot(pik3ca_boot,
                 aes(y = Treatment, x = `Point HR`, color = Robust)) +
  geom_point(size = 3.5) +
  geom_errorbar(
    aes(xmin = `CI Lower`, xmax = `CI Upper`),
    width = 0.2,
    linewidth = 0.9,
    orientation = "y"
  ) +
  geom_vline(xintercept = 1, linetype = "dashed",
             color = "gray40", linewidth = 0.9) +
  scale_color_manual(
    values = c("YES" = "#d62728", "NO" = "#aec7e8"),
    labels = c("YES" = "Robust (CI excludes 1)", "NO" = "Not robust")
  ) +
  labs(
    title = "B. PIK3CA \u2014 Bootstrap Validation (250 Resamples)",
    x = "Interaction HR (95% Bootstrapped CI)",
    y = "Treatment Comparison",
    color = "Robustness"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title = element_text(face = "bold", size = 11),
    legend.position = "bottom",
    legend.text = element_text(size = 9),
    axis.text = element_text(size = 9),
    plot.margin = margin(t = 20, r = 10, b = 10, l = 10)
  )

library(patchwork)
p_landmark + p_boot +
  plot_layout(widths = c(1.2, 1)) +
  plot_annotation(
    title = "Figure 3. PIK3CA \u00d7 Chemo+Targeted: Landmark Analysis and Bootstrap Validation",
    caption = paste0("Left: Interaction HRs by landmark period; * = FDR p < 0.05.\n",
                     "Right: Bootstrap 95% CIs from 250 resamples."),
    theme = theme(
      plot.title = element_text(face = "bold", size = 12),
      plot.caption = element_text(size = 8, color = "gray40")
    )
  )







#| fig-width: 14
#| fig-height: 8

library(broom)

survDF_pik3ca <- survDF |>
  mutate(
    PIK3CA_Status = ifelse(PIK3CA_mut == 1, "Mutant", "Wildtype"),
    Group = paste0("PIK3CA ", PIK3CA_Status, "\n", moa_consolidated)
  )

fit_pik3ca <- survfit(Surv(OS_time, OS_event) ~ Group, data = survDF_pik3ca)

# tidy the survfit object
km_df <- tidy(fit_pik3ca) |>
  mutate(
    strata = str_remove(strata, "Group="),
    strata = factor(strata)
  )

# color palette for 6 groups
pal6 <- c(
  "PIK3CA Mutant\nChemo"          = "#d62728",
  "PIK3CA Mutant\nChemo+Targeted" = "#ff7f0e",
  "PIK3CA Mutant\nOther/IO"       = "#9467bd",
  "PIK3CA Wildtype\nChemo"        = "#1f77b4",
  "PIK3CA Wildtype\nChemo+Targeted" = "#2ca02c",
  "PIK3CA Wildtype\nOther/IO"     = "#8c564b"
)

# clean labels for legend (replace \n with space)
km_df <- km_df |>
  mutate(strata_label = str_replace(strata, "\n", " | "))

pal6_clean <- setNames(pal6, str_replace(names(pal6), "\n", " | "))

p_pik3ca_manual <- ggplot(km_df, aes(x = time, y = estimate,
                                      color = strata_label,
                                      fill = strata_label)) +
  geom_step(linewidth = 0.9) +
  geom_ribbon(
    aes(ymin = conf.low, ymax = conf.high),
    alpha = 0.12, color = NA
  ) +
  scale_color_manual(values = pal6_clean, name = "Group") +
  scale_fill_manual(values = pal6_clean, name = "Group") +
  scale_x_continuous(
    breaks = seq(0, 400, by = 50),
    limits = c(0, 420)
  ) +
  scale_y_continuous(
    limits = c(0, 1),
    breaks = seq(0, 1, by = 0.25),
    labels = c("0.00", "0.25", "0.50", "0.75", "1.00")
  ) +
  labs(
    title = "Kaplan-Meier: Overall Survival by PIK3CA Mutation Status and Treatment",
    x = "Months from Treatment Start",
    y = "Overall Survival Probability"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", size = 12),
    legend.position = "right",
    legend.title = element_text(size = 10, face = "bold"),
    legend.text = element_text(size = 9),
    legend.key.height = unit(1.2, "cm"),
    legend.spacing.y = unit(0.3, "cm"),
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 11),
    panel.grid.minor = element_blank(),
    plot.margin = margin(t = 10, r = 10, b = 10, l = 10)
  ) +
  guides(
    color = guide_legend(override.aes = list(linewidth = 2)),
    fill = "none"
  )

print(p_pik3ca_manual)
```